---
title: "SP500MTH"
author: "Bo Suzow"
date: "March 17, 2018"
output: 
   html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

   library(readxl)  #http://readxl.tidyverse.org/
   library(dplyr)
   library(data.table)
   library(ggplot2)
   library(plotly)

   source('SP500MTH.R')

   
```

## Introduction

Since February 2018, volatility observed in the stock market makes average investors concerned. While the mean of SP500's calendar year returns (including dividends) since 1928 is 13.87%  (source)[http://www.stern.nyu.edu/~adamodar/pc/datasets/histretSP.xls], it took 13 years to recover to the 1999's peak from the two crashes.  In this report, we analyze rolling annualized returns to better understand volatilities which the calendar year based return averages may hide.
[Reference](https://www.thebalance.com/rolling-index-returns-1973-mid-2009-4061795)


***
## Data ETL
The monthly prices table hosted at [the data source](http://www.multpl.com/inflation-adjusted-s-p-500/table/by-month) has been extracted and loaded onto an Excel sheet.  The sheet is named SP500MTH.xls and stored in the local working directory. 

```{r}

sp500mth = read_excel("SP500mth.xls",sheet="sp500", range="A1:B1085")
sp500mth = sp500mth[-1,]  # remove the first row (which represents the index price of March 16.)
names(sp500mth) = c("mth","index")
sp500mth = as.data.frame(sp500mth[order(sp500mth$mth),])  # re-sort to make it ascending

# Refine the data frame, sp500mth

AR.df = annualRtn(sp500mth$index,sp500mth$mth)
sp500mth = sp500mth %>% 
               mutate(s.mth=AR.df$s.mth,AR = AR.df$AR) %>%
               
               mutate(class = as.factor(ifelse(AR>0,"pos","neg"))) %>%
               mutate(s.mth.10yr = AR.df$s.mth.10yr,
                      AR10=AR.df$AR10) %>% # adding the rolling annualized rtns
               mutate(class10yr = as.factor(ifelse(AR10>0,"pos","neg"))) %>%
               
               mutate(s.mth.20yr = AR.df$s.mth.20yr,
                      AR20=AR.df$AR20) %>% # adding the rolling annualized rtns
               mutate(class20yr = as.factor(ifelse(AR20>0,"pos","neg"))) %>%
   
               mutate(e.mth = as.Date(mth,origin="1970-01-01")) # change the mth column to the Data class
               

```

***

# Plot Title Attributes

```{r plottitle}

f <- list(
  family = "Arial",
  size = 18,
  color = "black"
)
x <- list(
  title = "Ending Month of the Annualized Rolling Return (ARR) Period",
  titlefont = f
)
y <- list(
  title = "Annualized Rolling Returns (ARR)",
  titlefont = f
)

```

# Plotting - Annualized Rolling Returns

```{r}

sp500ARR = sp500mth %>% select(s.mth,e.mth,AR,class)
names(sp500ARR) = c("s.mth","e.mth","ARR","class")
PlotARR(sp500ARR,year=1)

```

## Plotting 10-year ARR Averages

```{r}

sp500ARR = sp500mth %>% select(s.mth.10yr,e.mth,AR10,class10yr)
names(sp500ARR) = c("s.mth","e.mth","ARR","class")
PlotARR(sp500ARR,year=10)

```

## Plotting 20-year ARR Averages

```{r}

sp500ARR = sp500mth %>% select(s.mth.20yr,e.mth,AR20,class20yr)
names(sp500ARR) = c("s.mth","e.mth","ARR","class")
PlotARR(sp500ARR)
```

